OneD: increasing reproducibility of Hi-C Samples with abnormal karyotypes

### 2017-11-24 manuscript_nar.tex

Matrix-balancing methods (here \textit{ICE}, \textit{vanilla} and the
\textit{KR} algorithm) are specifically tailored for Hi-C because it is
the dominant 3C-derived technology. As a consequence, their performance is
typically lower when they are used for variants such as and Capture Hi-C
\citep{jager2015capture}. In this case, the contacts with one or more
regions of interest are enriched through direct purification, essentially
erasing all the other contacts from the Hi-C matrix. The lack of
homogeneity of the signal in such technologies makes normalization
particularly challenging.

We reasoned that \textit{OneD} should handle such cases gracefully,
because the signal is formally equivalent to an aberrant karyotype where
most of the genome is present in zero copy. We thus evaluated the
performance of \textit{OneD} on a Capture Hi-C data set at 5~Kbp
resolution centered on a 6~Mbp domain of chromosome 6 in T47D cells. We
found that the reproducibility score between replicates was higher after
normalization with \textit{OneD} than with matrix-balancing methods
(figure~\ref{fig:chic}A). The performance of \textit{OneD} was only
slightly lower than the raw data, suggesting that it barely disturbs the
biological signal. Other tools were not included beaus they either
crashed (\textit{caICB}) or had prohibitive run time (\textit{LGF}).

>>>>

Matrix-balancing methods (here \textit{ICE}, \textit{vanilla} and
\textit{KR}) are specifically tailored for Hi-C because it is the dominant
3C-derived technology. As a consequence, their performance is typically
lower when they are used for variants such as Capture Hi-C
\citep{jager2015capture}. In this case, the contacts with one or more
regions of interest are enriched through direct purification, essentially
erasing all the other contacts from the Hi-C matrix. The lack of
homogeneity of the signal in such technologies makes normalization
particularly challenging.

We reasoned that \textit{OneD} should handle such cases gracefully,
because the signal is formally equivalent to an aberrant karyotype where
most of the genome is present in zero copy. We thus evaluated the
performance of \textit{OneD} on a Capture Hi-C data set at 5~Kbp
resolution centered on a 6~Mbp domain of chromosome 6 in T47D cells. We
found that the reproducibility score between replicates was higher after
normalization with \textit{OneD} than with matrix-balancing methods
(figure~\ref{fig:chic}A). The performance of \textit{OneD} was only
slightly lower than the raw data, suggesting that it barely disturbs the
biological signal. Other tools were not included beause they had
prohibitive run time.
